PANCHANGE_data_preprocessing

### 2020-08-06 PANCHANGE_data_preprocessing2020-07-15.Rmd

Calculate sum score for GAD-7
Max should be 21.
```{r gad sum score prepan}
# Create vector with scoring key: If the item is reversed coded use -1. 
# There needs to be the correct number of values for each item in the questionnaire. 
gad.items_key_prepan <- c(1,1,1,1,1,1,1)

@@

Calculate sum score for PHQ-9
```{r phq sum score prepan}
# Create vector with scoring key: If the item is reversed coded use -1. 
# There needs to be the correct number of values for each item in the questionnaire. 
phq.items_key_prepan <- c(1,1,1,1,1,1,1,1,1)

@@

## COPING GLAD: GAD-7

@@

@@

##Missigness of questionnaire items
#PHQ - baseline
PHQ Missingness NAs per person - count and percentages
```{r PHQ basline missingness}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    na_per_person_phq_base =         # variable for NA per person
      rowSums(                       
        is.na(dat.no.sex.no.gender.no.age.no.ibd # sum the rows that are NA
          [,colnames(dat.no.sex.no.gender.no.age.no.ibd) #for all the rows and columns that are in the PHQ items at baseline
           %in%
             phq.items_base]
        )
    )
  )

@@

>>>>

Calculate sum score for GAD-7
Max should be 21.

*Note that scores with missingness will be changed to NA later in the script
```{r gad sum score prepan}
# Create vector with scoring key: If the item is reversed coded use -1. 
# There needs to be the correct number of values for each item in the questionnaire. 
gad.items_key_prepan <- c(1,1,1,1,1,1,1)

@@

Calculate sum score for PHQ-9

*Note that scores with missingness will be changed to NA later in the script
```{r phq sum score prepan}
# Create vector with scoring key: If the item is reversed coded use -1. 
# There needs to be the correct number of values for each item in the questionnaire. 
phq.items_key_prepan <- c(1,1,1,1,1,1,1,1,1)

@@

# glad ocir

```{r glad ocir cols}
glad.ocir.raw <- readRDS(file = paste0(data.raw_path, "/2020-08-03/glad/ocir_glad.rds"))
dim(glad.ocir.raw)
colnames(glad.ocir.raw)
```

```{r glad ocir cleaning}
glad.ocir.raw.id <- glad.ocir.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID),
         startDate.glad = startDate) %>% # Rename startDate
  select(
    Sample, # Sample
    ID, # ID
    ocir.i_repeatedly_check_doors_windows_drawers_etc_prepan = ocir.i_repeatedly_check_doors_windows_drawers_etc.,
    ocir.arranged_things_change_upset_prepan = ocir.arranged_things_change_upset,
    ocir.repeat_feel_numbers_prepan = ocir.repeat_feel_numbers,
    ocir.clean_simply_wash_feel_prepan = ocir.clean_simply_wash_feel,
    ocir.mind_upset_unpleasant_thoughts_prepan = ocir.mind_upset_unpleasant_thoughts,
    ocir.afraid_avoid_throwing_things_prepan = ocir.afraid_avoid_throwing_things,
    ocir.light_switches_water_taps_prepan = ocir.light_switches_water_taps,
    ocir.arranged_things_prepan = ocir.arranged_things,
    ocir.good_feel_bad_numbers_prepan = ocir.good_feel_bad_numbers,
    ocir.hands_longer_wash_prepan = ocir.hands_longer_wash,
    ocir.saved_things_prepan = ocir.saved_things,
    ocir.frequently_difficulty_rid_nasty_prepan = ocir.frequently_difficulty_rid_nasty,
    ocir.i_check_things_more_often_than_necessary_prepan = ocir.i_check_things_more_often_than_necessary.,
    ocir.arranged_properly_objects_upset_prepan = ocir.arranged_properly_objects_upset,
    ocir.feel_compelled_count_things_prepan = ocir.feel_compelled_count_things,
    ocir.strangers_touch_touched_difficult_prepan = ocir.strangers_touch_touched_difficult,
    ocir.control_difficult_find_thoughts_prepan = ocir.control_difficult_find_thoughts,
    ocir.i_collect_things_i_dont_need_prepan = ocir.i_collect_things_i_dont_need
  ) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(glad.ocir.raw.id)
# Inspect colnames
colnames(glad.ocir.raw.id)
#Differences
dim(glad.ocir.raw)[1]-dim(glad.ocir.raw.id)[1]
```

Calculate sum score for OCI-R

*Note that scores with missingness will be changed to NA later in the script
```{r ocir sum score prepan}
# Create vector with scoring key: If the item is reversed coded use -1. 
# There needs to be the correct number of values for each item in the questionnaire. 
ocir.items_key_prepan <- c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1)

# Create vector with items that are included in sum score
ocir.items_prepan = c(
  "ocir.i_repeatedly_check_doors_windows_drawers_etc_prepan",
  "ocir.arranged_things_change_upset_prepan",
  "ocir.repeat_feel_numbers_prepan",
  "ocir.clean_simply_wash_feel_prepan",
  "ocir.mind_upset_unpleasant_thoughts_prepan",
  "ocir.afraid_avoid_throwing_things_prepan",
  "ocir.light_switches_water_taps_prepan",
  "ocir.arranged_things_prepan",
  "ocir.good_feel_bad_numbers_prepan",
  "ocir.hands_longer_wash_prepan",
  "ocir.saved_things_prepan",
  "ocir.frequently_difficulty_rid_nasty_prepan",
  "ocir.i_check_things_more_often_than_necessary_prepan",
  "ocir.arranged_properly_objects_upset_prepan",
  "ocir.feel_compelled_count_things_prepan",
  "ocir.strangers_touch_touched_difficult_prepan",
  "ocir.control_difficult_find_thoughts_prepan",
  "ocir.i_collect_things_i_dont_need_prepan"
  )

# Calulate sum score
ocir.scored_items_prepan <- scoreItems(keys = ocir.items_key_prepan,
                                        items = glad.ocir.raw.id[ocir.items_prepan],
                                        totals = TRUE, 
                                        missing = TRUE, 
                                        impute = 'none', 
                                        min = 0, 
                                        max = 4)
# Add in column to tibble
glad.ocir.raw.id$ocir.sum_score_prepan <- ocir.scored_items_prepan$scores

# Look at values
glad.ocir.raw.id %>%
  descr(ocir.sum_score_prepan,
        stats = "common"
        )
```

```{r glad ocir head}
head(glad.ocir.raw.id)
```

```{r glad ocir summary}
summarytools::dfSummary(
  glad.ocir.raw.id,
  graph.col = F,
  valid.col = F
  )
```

# COPING GLAD: GAD-7

@@

# glad COPING ocir

```{r coping glad ocir cols}
coping.glad.ocir.raw <- readRDS(file = paste0(data.raw_path, "/2020-08-04/glad/ocir_coping_glad.rds"))
dim(coping.glad.ocir.raw)
colnames(coping.glad.ocir.raw)
```

```{r coping glad ocir summary labels}
summarytools::dfSummary(
  coping.glad.ocir.raw,
  graph.col = F,
  valid.col = F
  #labels.col = F
)
```

*Note that the COPING GLAD OCI-R ID numbers are stored under recipientFirstName (rather than ExternalDataReference)
```{r coping glad ocir cleaning}
coping.glad.ocir.raw.id <- coping.glad.ocir.raw %>%
  drop_na(recipientFirstName) %>% # Drop NAs
  distinct(recipientFirstName, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(recipientFirstName, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    ocir.i_repeatedly_check_doors_windows_drawers_etc_retro = ocir.i_repeatedly_check_doors_windows_drawers_etc.,
    ocir.arranged_things_change_upset_retro = ocir.arranged_things_change_upset,
    ocir.repeat_feel_numbers_retro = ocir.numbers_repeat_feel, ## Different from GLAD prepandemic
    ocir.clean_simply_wash_feel_retro = ocir.feel_contaminated_clean_simply, ## Different from GLAD prepandemic
    ocir.mind_upset_unpleasant_thoughts_retro = ocir.mind_upset_unpleasant_thoughts,
    ocir.afraid_avoid_throwing_things_retro = ocir.afraid_avoid_throwing_things,
    ocir.light_switches_water_taps_retro = ocir.light_switches_water_taps,
    ocir.arranged_things_retro = ocir.arranged_things,
    ocir.good_feel_bad_numbers_retro = ocir.bad_numbers_good_feel, ## Different from GLAD prepandemic
    ocir.hands_longer_wash_retro = ocir.longer_hands_wash, ## Different from GLAD prepandemic
    ocir.saved_things_retro = ocir.saved_things,
    ocir.frequently_difficulty_rid_nasty_retro = ocir.frequently_rid_difficulty_nasty, ## Different from GLAD prepandemic
    ocir.i_check_things_more_often_than_necessary_retro = ocir.i_check_things_more_often_than_necessary.,
    ocir.arranged_properly_objects_upset_retro = ocir.arranged_properly_objects_upset,
    ocir.feel_compelled_count_things_retro = ocir.feel_compelled_count_things,
    ocir.strangers_touch_touched_difficult_retro = ocir.strangers_touch_touched_object, ## Different from GLAD prepandemic
    ocir.control_difficult_find_thoughts_retro = ocir.control_find_difficult_thoughts, ## Different from GLAD prepandemic
    ocir.i_collect_things_i_dont_need_retro = ocir.i_collect_things_i_dont_need., ## Different from GLAD prepandemic
    ocir.i_repeatedly_check_doors_windows_drawers_etc_base = ocir.i_repeatedly_check_doors_windows_drawers_etc..1,
    ocir.arranged_things_change_upset_base = ocir.arranged_things_change_upset.1,
    ocir.repeat_feel_numbers_base = ocir.numbers_repeat_feel.1, ## Different from GLAD prepandemic
    ocir.clean_simply_wash_feel_base = ocir.feel_contaminated_clean_simply.1, ## Different from GLAD prepandemic
    ocir.mind_upset_unpleasant_thoughts_base = ocir.mind_upset_unpleasant_thoughts.1,
    ocir.afraid_avoid_throwing_things_base = ocir.afraid_avoid_throwing_things.1,
    ocir.light_switches_water_taps_base = ocir.light_switches_water_taps.1,
    ocir.arranged_things_base = ocir.arranged_things.1,
    ocir.good_feel_bad_numbers_base = ocir.bad_numbers_good_feel.1, ## Different from GLAD prepandemic
    ocir.hands_longer_wash_base = ocir.longer_hands_wash.1, ## Different from GLAD prepandemic
    ocir.saved_things_base = ocir.saved_things.1,
    ocir.frequently_difficulty_rid_nasty_base = ocir.frequently_rid_difficulty_nasty.1, ## Different from GLAD prepandemic
    ocir.i_check_things_more_often_than_necessary_base = ocir.i_check_things_more_often_than_necessary..1,
    ocir.arranged_properly_objects_upset_base = ocir.arranged_properly_objects_upset.1,
    ocir.feel_compelled_count_things_base = ocir.feel_compelled_count_things.1,
    ocir.strangers_touch_touched_difficult_base = ocir.strangers_touch_touched_object.1, ## Different from GLAD prepandemic
    ocir.control_difficult_find_thoughts_base = ocir.control_find_difficult_thoughts.1, ## Different from GLAD prepandemic
    ocir.i_collect_things_i_dont_need_base = ocir.i_collect_things_i_dont_need..1

# Inspect dimensions
dim(coping.glad.ocir.raw.id)
# Inspect colnames
colnames(coping.glad.ocir.raw.id)
#Differences
dim(coping.glad.ocir.raw)[1]-dim(coping.glad.ocir.raw.id)[1]
```

Calculate sum score for OCI-R retro
```{r coping ocir sum score retro}
# Create vector with scoring key: If the item is reversed coded use -1. 
# There needs to be the correct number of values for each item in the questionnaire. 
ocir.items_key_retro <- c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1)

# Create vector with items that are included in total score
ocir.items_retro = c(
  "ocir.i_repeatedly_check_doors_windows_drawers_etc_retro",
  "ocir.arranged_things_change_upset_retro",
  "ocir.repeat_feel_numbers_retro",
  "ocir.clean_simply_wash_feel_retro",
  "ocir.mind_upset_unpleasant_thoughts_retro",
  "ocir.afraid_avoid_throwing_things_retro",
  "ocir.light_switches_water_taps_retro",
  "ocir.arranged_things_retro",
  "ocir.good_feel_bad_numbers_retro",
  "ocir.hands_longer_wash_retro",
  "ocir.saved_things_retro",
  "ocir.frequently_difficulty_rid_nasty_retro",
  "ocir.i_check_things_more_often_than_necessary_retro",
  "ocir.arranged_properly_objects_upset_retro",
  "ocir.feel_compelled_count_things_retro",
  "ocir.strangers_touch_touched_difficult_retro",
  "ocir.control_difficult_find_thoughts_retro",
  "ocir.i_collect_things_i_dont_need_retro"
)

# Calulate total score
ocir.scored_items_retro <- scoreItems(keys = ocir.items_key_retro,
                                     items = coping.glad.ocir.raw.id[ocir.items_retro],
                                     totals = TRUE, 
                                     missing = TRUE, 
                                     impute = 'none', 
                                     min = 0, 
                                     max = 4)
# Add in column to tibble
coping.glad.ocir.raw.id$ocir.sum_score_retro <- ocir.scored_items_retro$scores

# Look at values
coping.glad.ocir.raw.id %>%
  descr(ocir.sum_score_retro,
        stats = "common"
  )
```

Calculate sum score for OCI-R base
```{r coping ocir sum score base}
# Create vector with scoring key: If the item is reversed coded use -1. 
# There needs to be the correct number of values for each item in the questionnaire. 
ocir.items_key_base <- c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1)

# Create vector with items that are included in total score
ocir.items_base = c(
  "ocir.i_repeatedly_check_doors_windows_drawers_etc_base",
  "ocir.arranged_things_change_upset_base",
  "ocir.repeat_feel_numbers_base",
  "ocir.clean_simply_wash_feel_base",
  "ocir.mind_upset_unpleasant_thoughts_base",
  "ocir.afraid_avoid_throwing_things_base",
  "ocir.light_switches_water_taps_base",
  "ocir.arranged_things_base",
  "ocir.good_feel_bad_numbers_base",
  "ocir.hands_longer_wash_base",
  "ocir.saved_things_base",
  "ocir.frequently_difficulty_rid_nasty_base",
  "ocir.i_check_things_more_often_than_necessary_base",
  "ocir.arranged_properly_objects_upset_base",
  "ocir.feel_compelled_count_things_base",
  "ocir.strangers_touch_touched_difficult_base",
  "ocir.control_difficult_find_thoughts_base",
  "ocir.i_collect_things_i_dont_need_base"
)

# Calulate total score
ocir.scored_items_base <- scoreItems(keys = ocir.items_key_base,
                                    items = coping.glad.ocir.raw.id[ocir.items_base],
                                    totals = TRUE, 
                                    missing = TRUE, 
                                    impute = 'none', 
                                    min = 0, 
                                    max = 4)
# Add in column to tibble
coping.glad.ocir.raw.id$ocir.sum_score_base <- ocir.scored_items_base$scores

# Look at values
coping.glad.ocir.raw.id %>%
  descr(ocir.sum_score_base,
        stats = "common"
  )
```

```{r coping glad ocir head}
head(coping.glad.ocir.raw.id)
```

```{r coping glad ocir summary}
summarytools::dfSummary(
  coping.glad.ocir.raw.id,
  graph.col = F,
  valid.col = F,
  labels.col = F)
```

@@

##Missingness of questionnaire items
#PHQ - baseline
PHQ Missingness NAs per person - count and percentages
```{r PHQ baseline missingness}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    na_per_person_phq_base =         # variable for NA per person
      rowSums(                       
        is.na(dat.no.sex.no.gender.no.age.no.ibd # sum the rows that are NA
          [,colnames(dat.no.sex.no.gender.no.age.no.ibd) #for all the rows and columns that are in the PHQ items at baseline
           %in%
             phq.items_base]
        )
    )
  )

@@

#OCI-R - baseline
OCI-R Missingness NAs per person - count and percentages
```{r OCI-R baseline missingness}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    na_per_person_ocir_base =         # variable for NA per person
      rowSums(                       
        is.na(dat.no.sex.no.gender.no.age.no.ibd # sum the rows that are NA
          [,colnames(dat.no.sex.no.gender.no.age.no.ibd) #for all the rows and columns that are in the OCI-R items at baseline
           %in%
             ocir.items_base]
        )
    )
  )

freq(dat.no.sex.no.gender.no.age.no.ibd$na_per_person_ocir_base) #have a look at the frequency

dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>% #NA per person as a percentage
  mutate(
    miss_pc_total_ocir_base = 
      na_per_person_ocir_base/18 # NA per person divided by the amount of items
  )

freq(dat.no.sex.no.gender.no.age.no.ibd$miss_pc_total_ocir_base)
```

#OCI-R - retrospective
OCI-R Missingness NAs per person - count and percentages
```{r OCI-R retrospective missingness}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    na_per_person_ocir_retro =         # variable for NA per person
      rowSums(                       
        is.na(dat.no.sex.no.gender.no.age.no.ibd # sum the rows that are NA
          [,colnames(dat.no.sex.no.gender.no.age.no.ibd) #for all the rows and columns that are in the OCI-R items at retroline
           %in%
             ocir.items_retro]
        )
    )
  )

freq(dat.no.sex.no.gender.no.age.no.ibd$na_per_person_ocir_retro) #have a look at the frequency

dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>% #NA per person as a percentage
  mutate(
    miss_pc_total_ocir_retro = 
      na_per_person_ocir_retro/18 # NA per person divided by the amount of items
  )

freq(dat.no.sex.no.gender.no.age.no.ibd$miss_pc_total_ocir_retro)
```

#OCI-R - prepandemic
OCI-R Missingness NAs per person - count and percentages
```{r OCI-R prepan missingness}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    na_per_person_ocir_prepan =         # variable for NA per person
      rowSums(                       
        is.na(dat.no.sex.no.gender.no.age.no.ibd # sum the rows that are NA
          [,colnames(dat.no.sex.no.gender.no.age.no.ibd) #for all the rows and columns that are in the OCI-R items at prepanline
           %in%
             ocir.items_prepan]
        )
    )
  )

freq(dat.no.sex.no.gender.no.age.no.ibd$na_per_person_ocir_prepan) #have a look at the frequency

dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>% #NA per person as a percentage
  mutate(
    miss_pc_total_ocir_prepan = 
      na_per_person_ocir_prepan/18 # NA per person divided by the amount of items
  )

freq(dat.no.sex.no.gender.no.age.no.ibd$miss_pc_total_ocir_prepan)
```

Count for OCI-R baseline, retrospective, and prepandemic missingness per person
```{r count of people missing a number of items for OCI-R}
#baseline
table(dat.no.sex.no.gender.no.age.no.ibd$na_per_person_ocir_base)
#retrospective
table(dat.no.sex.no.gender.no.age.no.ibd$na_per_person_ocir_retro)
#pre-pandemic
table(dat.no.sex.no.gender.no.age.no.ibd$na_per_person_ocir_prepan)
```
